---
title: "Read a data product"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read a data product}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(daapr)
```

# Background

daapr provides a minimalist and uniform interface for retrieving
data products and the associated metadata no matter where the data product is
housed. Currently, s3, local, and LabKey are three types of remote repositories for
housing data products.

This vignette covers reading a data product that already exists. To build and
deploy one, see the
[new project workflow](https://daapr-team.github.io/daapr/articles/daapr.html).

daapr leverages the `pins` package for accessing data. As such, it
leverages the concept of `boards` for organizing data into **data spaces**. As
an example, a team working on multiple clinical studies, may consider defining 
a data space or `board` for each study, each of which defined by a governance 
policy. In this example, a `board` is not dissimilar to the 
idea of a directory or folder. A board can be instantiated on different data 
platforms (e.g. AWS s3) supported by `pins` or its extensions. 

As different data platforms (e.g. s3 vs. LabKey) have their own specific url and
ways of authentication, different parameters need to be set up for how 
connection to the remote data repository is to be established. Once the 
connection is established, however, all other interactions (e.g. list, get, 
version, etc.) are done using the same set of functions.  With that in mind, 
the steps outlined below are divided into 

1. Setting up the connection
2. Interacting with the data product

# Set up connection

Setting up connection to data products involves setting

1. Board parameters: parameters relating to the remote location of the data 
product
2. Credential parameters: parameters checked by the remote location of the data 
when trying to access

**NOTE:** prior to proceeding further verify that you have working credentials 
for the repository you need to access.


## Board parameters

- For s3, you need to know the bucket name and the region.
- For LabKey, you need to know the url of platform and specific folder.
- For local boards, you only need to provide the path of the local folder


## Format board parameters

For S3, you need to provide the s3 bucket name and region

```{r, eval=FALSE}
board_params <- board_params_set_s3(
  bucket_name = "bucket_name",
  region = "us-east-1"
)
```

If the data product is stored under a prefix within the bucket rather than at
the bucket root, pass that prefix as well. The prefix must end with a trailing
slash.

```{r, eval=FALSE}
board_params <- board_params_set_s3(
  bucket_name = "bucket_name",
  prefix = "data-products/",
  region = "us-east-1"
)
```

For local, you need to provide the absolute folder path

```{r, eval=FALSE}
board_params <- board_params_set_local(
  folder = "/project_folder/subfolder"
)
```

For LabKey, you need to provide the server url and the path of the study folder
on that server. A `cache_alias` can optionally be given to name the local cache
folder.

```{r, eval=FALSE}
board_params <- board_params_set_labkey(
  url = "https://learn.labkey.com/",
  folder = "folder_name"
)
```


## Format your credentials

This can be done with `cred_set_aws` or `cred_set_labkey` functions. The 
cred_set function signatures informs you of the type of parameters needed.

**NOTE:** For security reasons, avoid putting your credentials inside your 
script or inside any files that are version-controlled. It is highly recommended 
to establish a habit of using environment variables in conjunction with secret 
management packages like `keyring`. Credentials can be set as an environment variable
via an `.Renviron` file or via a secret manager package such as `keyring`. 
For AWS credentials, users can also store these through standard aws configuration files
and load them into their session using the `aws.signature` package.

The signature for creds for AWS s3 is takes two parameters.

```{r, eval=FALSE}
aws_creds <- creds_set_aws(
  key = Sys.getenv("AWS_ACCESS_KEY_ID"),
  secret = Sys.getenv("AWS_SECRET_ACCESS_KEY")
)
```

For LabKey, a single API key is needed. See the
[LabKey documentation on API keys](https://www.labkey.org/Documentation/wiki-page.view?name=apikey)
for how to generate one.

```{r, eval=FALSE}
labkey_creds <- creds_set_labkey(api_key = Sys.getenv("LABKEY_API_KEY"))
```

Local boards need no credentials, so `creds` can be omitted when connecting to
one.


# Connect and interact with data product

With the connection parameters established, interaction with the data product 
use the same set of function and format no matter where the data is housed.

## Connect to the board

Then we can connect to the board.

```{r, eval=FALSE}
board_object <- dp_connect(board_params = board_params, creds = aws_creds)
```

## Work with the board

List what data products available on this board we just connected to. Note we 
need to pass the board object in order to list available data products. 

```{r, eval=FALSE}
dp_list(board_object = board_object)
```

## Retrieve data

Load data into the working environment, e.g. as object `dp`. Specifying version is optional. 

```{r, eval=FALSE}
dp <- dp_get(
  board_object = board_object,
  data_name = "dp-studyid_branchid",
  version = "c5a51c3"
)
```
